---
title: "R Notebook"
output: html_notebook
---

#### This code is a snapshot of the complete code presented in the file "ranking.methods". It seeks to show more easily the split between training and testing datasets and how it is applied to the 2 types of approach: dynamic and static. 

###libraries
```{r}
library(ggplot2)
library(directlabels)
library(compete)
devtools::install_github("tbonne/rankReliability", force=TRUE)
library(rankReliability)
library(lubridate)
library(plyr)
library(dplyr)
library(EloRating)
```

### Load needed functions
```{r}
## CREATE DATAFRAME FUNCTION
create.dataframe<- function(training.data){
  
  List<- training.data[,3:4]
  Individuals<-c(as.character(List$from),as.character(List$to))
  
  #Delete replications
  Individual.vector<-unique(Individuals)
  nb.individual <- length(Individual.vector)
  
  #create a data frame out of a transposed vector
  Obs.Frame = as.data.frame(t(Individual.vector))
  #change the names of the dataframe to be IDs
  colnames(Obs.Frame) <- Individual.vector # The outcome isn't the best as the fiRBM row contains all the IDs again, but it's an easy fix (we will delete the fiRBM row later)
  
  return(Obs.Frame)
}

## EDGELIST FUNCTION
create.an.edgeList<-function(nn){
  
  Edgelist<-dplyr::count(nn, winner , loser)
  
  return(Edgelist)
}

rbind.all.columns <- function(x, y) {
  
  x.diff <- setdiff(colnames(x), colnames(y))
  y.diff <- setdiff(colnames(y), colnames(x))
  
  x[, c(as.character(y.diff))] <- NA
  
  y[, c(as.character(x.diff))] <- NA
  
  return(rbind(x, y))
}

```

### Import dataset. 
```{r}
## Import dominance data
Dominance <- read.csv ("dominance.data.csv")

##FYI: in the result column a win =1, a loss=2, a draw=3 and unknown result = 4
## Get rid of unknown outcomes.
Dominance.clear<-Dominance[-which(Dominance$result=="4"),]

## Get rid of individuals who appear in testing dataset only as we wont have any ranks associated to them from the training dataset.
Dominance.clear2<-Dominance.clear[-which (Dominance.clear$to=="macy"),]
Dominance.clear3<-Dominance.clear2[-which (Dominance.clear2$from=="macy"),]
Dominance.clear4<-Dominance.clear3[-which (Dominance.clear3$to=="rodr"),]
Dominance.clear5<-Dominance.clear4[-which (Dominance.clear4$from=="rodr"),]
Dominance.clear6<-Dominance.clear5[-which (Dominance.clear5$to=="balu"),]
Dominance.clear7<-Dominance.clear6[-which (Dominance.clear6$from=="balu"),]
Dominance.clear8<-Dominance.clear7[-which (Dominance.clear7$to=="nige"),]
## Set up date format and make sure interactions are chronologically ordered
Dominance.clear8$date <- lubridate::ymd(as.character(Dominance.clear8$date))
Dominance.df <- Dominance.clear8[order(Dominance.clear8$date),] 
```

###### PART 1 
#### Split the dataset into training and testing
```{r}
## here we want to keep the chronological order hence using the function filter and not just the sample function
# TRAINING dataset (80% OF THE DATA)
training.data<- Dominance.df %>% filter(date<="2017-04-25")

# TESTING dataset (20% OF THE DATA)
testing.data<- Dominance.df %>% filter(date>"2017-04-25")
```

#### Create main dataframe that will store the ranks
```{r}
## create main df to store ranks
main.df<- create.dataframe(training.data)
main.df<-cbind(main.df,Method = "method")
main.df$Method<-as.character(main.df$Method)
```

### First start by inferring and extracting the individual ranks, using the training dataset. We offer two examples: one using a dynamic approach (Elo-rating) and the second using a matrix-based approach (David's scores)

## ORIGINAL ELO RATING MEHTOD
```{r}
## Run elo on the TRAINING data ONLY
elo.scores <- elo.seq(winner=training.data$winner, loser=training.data$loser, Date=training.data$date,  runcheck=FALSE, draw =  training.data$draw)

# Select the scores for the latest date in the training dataset
latest.scores<-extract_elo(elo.scores, "2017-04-25") # NB: individuals who died during this period will have a "NA" score

## Store elo rating per individual
elo.df<-as.data.frame(t(sort(latest.scores)))
# transform ratings into ordinal numbers
elo.df[2,]<- c(62:1) 
elo.df<-elo.df[-1,]
```

## DAVID'S SCORES using the compete package and the method "p"
```{r}
## Isolate winner-loser to get a matrix with extra column giving the results
outcome.data<- training.data[,c("winner","loser","result")]
outcome.data$result[outcome.data$result %in% "1"]<-"W"
outcome.data$result[outcome.data$result %in% "2"]<-"L"
outcome.data$result[outcome.data$result %in% "3"]<-"T"

matrix.result<- get_wl_matrix(outcome.data, ties = "keep")

## Method P
David.score.p<-ds(matrix.result, norm = TRUE, type = "P")

## Store david's scores
dataframe<-as.data.frame (t(sort(David.score.p)))
#transform scores into ordinal numbers
dataframe[2,]<-c(62:1)
dataframe<-dataframe[-1,]
```

#compile elo and david's scores dataframes

```{r}
main.df<-as.data.frame(t(rbind.fill(list(elo.df, dataframe))))
colnames(main.df)<- c("elo","david")
main.df$ID<- rownames(main.df)
rownames(main.df)<-c()
rank.df<- main.df[,c(3,1,2)]

rank.df$elo<-as.integer(rank.df$elo)
rank.df$david<-as.integer(rank.df$david)
```

## Create dataframe to store data when looking at whether ranks match aggressive outcomes in the testing dataset

```{r}
Decay.dataframe<- as.data.frame(matrix(0, ncol = 2, nrow = nrow(testing.data)))
colnames(Decay.dataframe)<- c("Original.elo", "compete.p")

## df to store day.nb
Day.df<- as.data.frame(matrix(0, ncol = 2, nrow = nrow(testing.data)))
colnames(Day.df)<- c("Original.elo", "compete.p")
```

## use the rankReliability package to look at the percentage of reliability 

```{r}
relia.res<- reliability_check(rank.df, testing.data[,c(1,8,9,5,10)])

reliability<- relia.res [[1]]
match.reliability<- relia.res[[2]]

# Plot the results
plots.res<-reliability_plot(reliability,match.reliability)
```

###### PART 2: modify the training dataset length and calculate ranks for each length of it

#####Isolate individual ID that we will bind to the matching ranks
```{r}
List<- training.data[,3:4]
Individuals<-c(as.character(List$from),as.character(List$to))

Individual.vector<-unique(Individuals)
df.ind<- data.frame(Individual.vector)
colnames(df.ind)<- c("ID")
```

### Run a loop to extract ranks for each different training dataset length. Loop done for each tested method.

## ELO-RATING ORIGINAL
```{r}
## For my own simplicity, in what follows i use the number of days (instead of the date column).

df.elo<- create.dataframe(training.data)
Day<-vector()

for (m in seq(6,786, by=60)) {
  
  windowStart<-m
  sub.df<-  training.data %>% filter(day_nb>=m)
  
  ## Run elo: avec draw AND presence included. 
  res<- elo.seq(winner=as.character(sub.df$winner), loser=as.character(sub.df$loser), Date=sub.df$date,runcheck=FALSE, draw =  sub.df$draw)
  
  ## Select the scores for the latest date
  latest.scores<-extract_elo(res)
  
  ## Store elo ratings
  dataframe<-as.data.frame(t(sort(latest.scores)))
  dataframe[2,]<- c(length(latest.scores):1)
  dataframe<-dataframe[-1,]
  df.elo<-rbind.fill(list(df.elo, dataframe))
  Day[length(Day)+1]<-windowStart
  
}

#store ranks
df.elo<- df.elo[-1,]
Elo.ranks<- cbind(df.elo, Day)
Elo.ranks<- as.data.frame(t(Elo.ranks))
colnames(Elo.ranks)<-as.character(unlist(Elo.ranks[63,]))
Elo.ranks<-Elo.ranks[-63,]
Elo.ranks[,15]<- c("Elo.original")
names(Elo.ranks)[length(names(Elo.ranks))]<-"Method" 
Elo.ranks<-cbind(Elo.ranks,df.ind)
```

## DAVID'S SCORE
## Compete package
```{r}
Day<-vector()
main.df.p<- create.dataframe(training.data)

# Run loop
for (m in seq(6,786, by=60)) {
  
  windowStart<-m
  sub.df<-  training.data %>% filter(day_nb>=m)
  
  ## Isolate winner-loser to get a matrix with extra column giving the results
  outcome.data<- sub.df[,c("winner","loser","result")]
  outcome.data$result[outcome.data$result %in% "1"]<-"W"
  outcome.data$result[outcome.data$result %in% "2"]<-"L"
  outcome.data$result[outcome.data$result %in% "3"]<-"T"
  
  matrix<- get_wl_matrix(outcome.data, ties = "keep")
  
  ## Get rank (DS), Type P
  David.score.p<-ds(matrix, norm = TRUE, type = "P")
  dataframe<-as.data.frame(t(sort(David.score.p)))
  dataframe[2,]<- c(length(David.score.p):1)
  dataframe<-dataframe[-1,]
  main.df.p<-rbind.fill(list(main.df.p, dataframe))
  
  Day[length(Day)+1]<-windowStart
  
}

#Store ranks
main.df.p<- main.df.p[-1,]
DS.ranks.P<- cbind(main.df.p, Day)
DS.ranks.P<- as.data.frame(t(DS.ranks.P))
colnames(DS.ranks.P)<-as.character(unlist(DS.ranks.P[63,]))
DS.ranks.P<-DS.ranks.P[-63,]
DS.ranks.P[,15]<- c("DS.compete.P")
names(DS.ranks.P)[length(names(DS.ranks.P))]<-"Method" 

Ranks.compete.P<-cbind(DS.ranks.P,df.ind)
```

## compile both dataframes
```{r}
rank.training<- rbind(Ranks.compete.P,Elo.ranks)
rownames(rank.training)<-c()
rank.training<-rank.training[,c(16,15,1:14)]
```


##test methods's reliability: do ranks match aggressive outcomes in the testing dataset?
```{r}
## looking at David's scores reliability through time.
rank.df<-rank.training%>%filter(Method=="DS.compete.P") # select for this method
rank.df<-rank.df[,-2]

david.relia.res<-reliability_check(rank.df,testing.data[,c(1,8,9,5,10)])
relia.df<-david.relia.res[[1]]

relia.df<-relia.df %>% mutate(Month=as.numeric(sub('Match_', '', as.character(relia.df$Method))))
relia.df$Method<- "David's.scores"

ggplot(relia.df, aes(x=Month, y=Reliability))+ geom_line()

## looking at the Elo rating reliability through time.
rank.df.elo<-rank.training%>%filter(Method=="Elo.original") # select for this method
rank.df.elo<-rank.df.elo[,-2]

elo.relia.res<-reliability_check(rank.df.elo,testing.data[,c(1,8,9,5,10)])
relia.df.elo<-elo.relia.res[[1]]

relia.df.elo<-relia.df.elo %>% mutate(Month=as.numeric(sub('Match_', '', as.character(relia.df.elo$Method))))
relia.df.elo$Method<- "Elo.original"

## Combine the 2 dataframes
df.methods <- rbind(relia.df,relia.df.elo)


#Plot
ggplot(df.methods, aes(x=Month, y=Reliability,color = Method))+ geom_line()
```

###### PART 3: keep the training dataset constant and modify the testing dataset length

####  New split of training and testing datasets. Here we split sooner so we have a much bigger testing dataset to play around. 
```{r}
## Split data
# TRAINING dataset
newtraining.data<- Dominance.df %>% filter(date<="2015-07-04")

# TESTING dataset
newtesting.data<- Dominance.df %>% filter(date>"2015-07-04")
```

## Obtain dataframe of individual IDs
```{r}
List<- newtraining.data[,3:4]
Individuals<-c(as.character(List$from),as.character(List$to))

Individual.vector<-unique(Individuals)
df.ind<- data.frame(Individual.vector)
colnames(df.ind)<- c("ID")
```


### Extract ranks with each method using the new training dataset
## ORIGINAL ELO-RATING
```{r}
df.elo<- create.dataframe(newtraining.data)

result.elo <- elo.seq(winner=as.character(newtraining.data$winner), loser=as.character(newtraining.data$loser), Date=newtraining.data$date,runcheck=FALSE, draw =  newtraining.data$draw)

## Select the scores for the latest date
latest.scores<-extract_elo(result.elo)

## Store mother's elo ratings
## Store elo rating per individual
elo.df<-as.data.frame(t(sort(latest.scores)))
elo.df[2,]<- c(43:1)
elo.df<-elo.df[-1,]
rownames(elo.df)<-c()
```

## DAVID'S SCORES
#Compete package
```{r}
## Create storing df
main.df.p<- create.dataframe(newtraining.data)

## Isolate winner-loser to get a matrix with extra column giving the results
outcome.data<- newtraining.data[,c("winner","loser","result")]
outcome.data$result[outcome.data$result %in% "1"]<-"W"
outcome.data$result[outcome.data$result %in% "2"]<-"L"
outcome.data$result[outcome.data$result %in% "3"]<-"T"

matrix<- get_wl_matrix(outcome.data, ties = "keep")

## Get rank (DS), Type P
David.score.p<-ds(matrix, norm = TRUE, type = "P")
## Store david's scores
dataframe<-as.data.frame (t(sort(David.score.p)))
dataframe[2,]<-c(43:1)
dataframe<-dataframe[-1,]
dataframe[2,]<-names(dataframe)
rownames(dataframe)<-c()
```

#compile elo and david's scores dataframes
```{r}
rank.df<-as.data.frame(t(rbind.fill(list(elo.df, dataframe))))
colnames(rank.df)<- c("elo","david","ID")
rank.df$elo<-as.integer(as.character(rank.df$elo))
rank.df$david<-as.integer(as.character(rank.df$david))

rank.df<-rank.df[,c(3,1,2)]
rownames(rank.df)<- c()
```

##  test if the rank orders match with dyadic outcomes from testing dataset
```{r}
## here have to make the size of the testing datset vary
start <- 186
end<-1086
windowsize<- 60

final.df <- data.frame(Method=as.character("elo"), Reliability=1,low.ci.boot=1,high.ci.boot=1,Month=1)

while (start + windowsize<=end) {
  
  m<-start + windowsize
  sub.prediction<-  newtesting.data %>% filter(day_nb<=m)
  # Calculate efficiency
  prediction.df<- reliability_check(rank.df, sub.prediction[,c(1,8,9,5,10)])
  
  reliability.df<- prediction.df[[1]]
  reliability.df$Month <- m #store month
  
  final.df<-rbind(final.df,reliability.df)
  windowsize<- windowsize+60
  
} 

final.df<- final.df[-1,]

#Plot
ggplot(final.df, aes(x=Month, y=Reliability,color = Method))+ geom_line()
```

## Group and dyad level changes in outcome predictability
```{r}
## Split data: keep only 80% of it. Will predict with 20% of it.
Dominance.clear$date<- lubridate::ymd(as.character(Dominance.clear$date))
Dominance.clear <- Dominance.clear[order(Dominance.clear$date),] 

train.df<-  Dominance.clear %>% filter(date<="2015-05-06")

###20% left used to predict
long.testing.df<-Dominance.clear %>% filter(date>"2015-05-06")
```

## get rank order: I&SI method
```{r}
outcome.data<- train.df[,c("winner","loser","result")]
outcome.data$result[outcome.data$result %in% "1"]<-"W"
outcome.data$result[outcome.data$result %in% "2"]<-"L"
outcome.data$result[outcome.data$result %in% "3"]<-"T"

matrix.result<- get_wl_matrix(outcome.data, ties = "keep")
isi.order13<-isi13(matrix.result, nTries = 450,random = FALSE)
df.finale <- data.frame(isi.order13$best_order)
df.finale$rank <- seq.int(nrow(df.finale))
colnames(df.finale)<- c("ID", "rank")
```

## get reliability checked
```{r}
## Calculate rank order reliability
Decay.ds<-reliability_check(df.finale,long.testing.df[,c(1,8,9,5,10)])

## store each df
df.reliability<-Decay.ds[[1]]
decay<-Decay.ds[[2]]


write.csv(decay, "decay.pre.dyad.csv")
## add dyad: for ease, i did it manually ... and loaded back the file that had each outcome associated with the dyad type
decay.isi<- read.csv("decay.dyad.csv")
decay.isi$date<- ymd(decay.isi$date)

# isolate dyadd: females
female.df<-decay.isi%>%filter(dyad=="Af_Af")
female.df<- female.df [,-c(1,8)]

# isolate dyadd: males
male.df<-decay.isi%>%filter(dyad=="Am_Am")
male.df<- male.df[,-c(1,8)]


decay.isi<-decay.isi[,-c(1,8)]
```

## Plots
```{r}
# Global trend
global<-reliability_plot(df.reliability,decay.isi,method="loess")
plot.global<- global[[2]]
plot.global+geom_smooth(method="loess",color="blue") 

# adult female dyads
female.dyad<-reliability_plot(df.reliability,female.df,method="loess")
plot.female<- female.dyad[[2]]
plot.female+geom_smooth(method="loess",color="blue")


# adult male dyads 
male.dyad<-reliability_plot(df.reliability,male.df,method="loess") 
plot.male<- male.dyad[[2]]
plot.male+geom_smooth(method="loess",color="blue")
```

